---
title: "R assessment 2"
output:
  pdf_document: default
  html_document: default
date: "2022-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pressure, echo=FALSE}
library(tidyverse)
library(corrplot)
library(GGally)
library(leaps)
library(ggfortify)
library(car)
```

```{r cars}
df <- read.csv("/Users/parkjs9416/Desktop/R_project2/Life Expectancy Data.csv")
```


```{r}
#Checking and Removing N/A
sum(is.na(df))
```
This dataset includes 2563 of N/A

Removing unneccessary columns which are year location and status
```{r}
df <- df[, -c(1:2)]
```

```{r}
df <- drop_na(df)
```

```{r}
sum(is.na(df))
```


```{r}
levels(df$Status) <- c(1,0)
#turned the status into the 0 and 1
```


```{r}
df$Status = as.numeric(df$Status=='Developing')
#turned the status into the 0 and 1 upper one not worked
#developing == 1, and data is numeric data
```
making a new data frame which has same variables like original data frame, it is because box-cox was not
possible with the original data frame, so new data frame but has same variables and obs were made
```{r}
new_data1 <- df %>% select(Life.expectancy,Status,Adult.Mortality,infant.deaths,
                           Alcohol,percentage.expenditure,Hepatitis.B,Measles,BMI,under.five.deaths,
                          Polio,Total.expenditure,Diphtheria,HIV.AIDS,GDP,thinness..1.19.years,
                          thinness.5.9.years,Income.composition.of.resources,Population,Schooling)
```

```{r}
#move Life.expectancy column behind of Status
df<- df %>% relocate(Status, .after = Life.expectancy)
```


Explore the data
```{r}
#produce appropriate visualisations and summary statistics which reflect the key relationships
df_life <- cor(new_data1)
life <- df_life[1,1:20]
life
```
```{r}
summary(new_data1)
```



```{r}
my_data1 <- new_data1[, c(1:11)]
my_data2 <-  new_data1[,c(1,12:20)]
```

```{r}
ggpairs(my_data1,
    upper = list(continuous = wrap("cor", size = 3)),
    lower = list(continuous = wrap("points", alpha = 0.3,    size=0.1)))+
  theme_grey(base_size = 5)
ggpairs(my_data1,
        upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = wrap("points", alpha = 0.3,    size=0.1)))+
  theme_grey(base_size = 5)
```


```{r}
ggcorr(new_data1, label = TRUE, label_round = 2,label_size = 2,hjust=1,
       layout.exp = 5,
       name="Correlation")+
  ggplot2::labs(title = "Correlation of each variable")


```

```{r}
#full model
full_mod <- lm(Life.expectancy~., data=new_data1)
vif(full_mod)
```
Infant.death and uner.five.death --> removed

```{r}
full_mod<-lm(Life.expectancy~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
             under.five.deaths+BMI+Income.composition.of.resources+percentage.expenditure+
             Polio+thinness..1.19.years+thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
             Total.expenditure+GDP+Schooling,data=new_data1)
vif(full_mod)
```
```{r}
full_mod<-lm(Life.expectancy~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
               under.five.deaths+BMI+Income.composition.of.resources+Polio+
               thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
               Total.expenditure+GDP+Schooling,data=new_data1)
vif(full_mod)
```

```{r}
new_data2 <- df %>% select(Life.expectancy,Status,Adult.Mortality,
                           Alcohol,Hepatitis.B,Measles,BMI,under.five.deaths,
                          Polio,Total.expenditure,Diphtheria,HIV.AIDS,GDP,
                          thinness.5.9.years,Income.composition.of.resources,Population,Schooling)
```

```{r}

#heatmap with new data2
ggcorr(new_data2, label = TRUE, label_round = 2,label_size = 2,hjust=1,
       layout.exp = 5,
       name="Correlation") + ggplot2::labs(title = "Correlation of after Multicollinearity")


```



```{r}
autoplot(full_mod)
```

```{r}
ggplot(new_data2) + geom_boxplot(aes(Life.expectancy ,fill = Life.expectancy)) 
```

The following histograms are for checking variables whether transformation is needed or not.
```{r}
#Checking variables which are needed to be transformed
hist(new_data2$Alcohol)
hist(sqrt(new_data2$Alcohol))
scatterplot(Life.expectancy~sqrt(new_data2$Adult.Mortality), data=new_data2)
hist(log1p(new_data2$Diphtheria))
hist(sqrt(new_data2$thinness.5.9.years))
hist(new_data2$Hepatitis.B)
hist(log1p(new_data2$Hepatitis.B))
hist(new_data2$Polio)
hist((new_data2$Polio)^1.5)
hist(new_data2$GDP)
hist(log(new_data2$GDP))
hist(new_data2$HIV.AIDS)
hist(log10(new_data2$HIV.AIDS))
hist(new_data2$Hepatitis.B)
hist(new_data2$Hepatitis.B)
hist(new_data2$Adult.Mortality)
hist(sqrt(new_data2$Adult.Mortality))
hist(new_data2$under.five.deaths)
hist(log(new_data2$under.five.deaths))
hist(new_data2$Total.expenditure)
hist(new_data2$Schooling)
hist(new_data2$Income.composition.of.resources)
```

```{r}
#Transform variables
hist(new_data2$Adult.Mortality)
hist(sqrt(new_data2$Adult.Mortality))

hist(new_data2$Alcohol)
hist(sqrt(new_data2$Alcohol))

hist(new_data2$GDP)
hist(log(new_data2$GDP))

hist(new_data2$thinness.5.9.years)
hist(log(new_data2$thinness.5.9.years))

hist(new_data2$Population)
hist(log(new_data2$Population))
```

```{r}
#Refelct transformed data
new_data2 %>% 
  mutate(Adult.Mortality=sqrt(Adult.Mortality),
         Alcohol=sqrt(Alcohol),
         GDP = log(GDP),
         thinness.5.9.years = log(thinness.5.9.years),
         Population = log(Population)) ->
  new_dataTrans
```

```{r}
sum(is.na(new_dataTrans))
```


###Model Building
```{r}
#leap and bound model building
result.life <- leaps(x=new_dataTrans[,2:17], y=new_dataTrans[,1],method = "adjr2",names=names(new_dataTrans)[-1])
fit <- data.frame(Size = result.life$size,
                  adjr2=round(result.life$adjr2, 3),
                  result.life$which, row.names=NULL)
ggplot(fit)+geom_point(aes(Size, adjr2))

```
Leap and Bound shows that models with 9, 10, 11 variables can be made and tested
```{r}
fit %>% filter(Size == 9)
fit %>% filter(Size == 10)
fit %>% filter(Size == 11)
```
when it is size 9, variables can be Life.expectancy,Adult.Mortality, percentage.expenditure, BMI, Diphtheria, HIV.AIDS, Income.composition.of.resources, Schooling
When it is size 10
When it is size 11
```{r}
size_9 <- lm(Life.expectancy ~ Adult.Mortality + Polio + Diphtheria+ HIV.AIDS+
               GDP+thinness.5.9.years+ Income.composition.of.resources+Schooling, data= new_dataTrans)

size_10 <- lm(Life.expectancy ~ Adult.Mortality + BMI + Polio + Diphtheria+HIV.AIDS+
                GDP+thinness.5.9.years+Income.composition.of.resources+Schooling,data=new_dataTrans)

size_11<- lm(Life.expectancy ~  Status+Adult.Mortality+BMI+Polio+Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
              Income.composition.of.resources+Schooling , data=new_dataTrans)
```


```{r}
#Stepwise selection

int_mod <- lm(Life.expectancy~ 1, data=new_dataTrans)
step(int_mod, scope=~Status+Adult.Mortality+HIV.AIDS+Diphtheria+BMI+Income.composition.of.resources+
       Polio+thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+Total.expenditure+
       GDP+under.five.deaths+Schooling, direction = "both")
```

```{r}
#stepwise mode
Step_mod <- lm(Life.expectancy ~ Schooling + HIV.AIDS + Adult.Mortality + 
    Diphtheria + thinness.5.9.years + Income.composition.of.resources + 
    GDP + Polio + BMI + Status + Measles + Hepatitis.B + Population, 
    data = new_dataTrans)

```


```{r} 
#Backward selection
full_mod <- lm(Life.expectancy ~ Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+under.five.deaths+
                 Polio + Total.expenditure+Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
                 Income.composition.of.resources+Population+Schooling, data=new_dataTrans)
drop1(full_mod, test="F",
      scope=~Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+under.five.deaths+
                 Polio + Total.expenditure+Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
                 Income.composition.of.resources+Population+Schooling)

```
```{r}
#Total expenditure is the least significant, refit without Total.expenditure
mod <- lm(Life.expectancy ~ Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+under.five.deaths+
            Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+Income.composition.of.resources+
            Population+Schooling, data=new_dataTrans)
drop1(full_mod, test="F",
      scope=~Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+under.five.deaths+
                 Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
                 Income.composition.of.resources+Population+Schooling)

```
```{r}
#under.five.deaths is the least significant so refit model without under.five.deaths
mod <- lm(Life.expectancy ~ Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+
            Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+Income.composition.of.resources+
            Population+Schooling, data=new_dataTrans)
drop1(full_mod, test="F",
      scope=~Status+Adult.Mortality+Alcohol+Hepatitis.B+Measles+BMI+
                 Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
                 Income.composition.of.resources+Population+Schooling)


```
```{r}
#now, alcohol is the least significant
mod <- lm(Life.expectancy ~ Status+Adult.Mortality+Hepatitis.B+Measles+BMI+
            Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+Income.composition.of.resources+
            Population+Schooling, data=new_dataTrans)
drop1(full_mod, test="F",
      scope=~Status+Adult.Mortality+Hepatitis.B+Measles+BMI+
                 Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+
                 Income.composition.of.resources+Population+Schooling)
```
```{r}
Back_mod <- lm(Life.expectancy ~ Status+Adult.Mortality+Hepatitis.B+Measles+BMI+
            Polio +Diphtheria+HIV.AIDS+GDP+thinness.5.9.years+Income.composition.of.resources+
            Population+Schooling, data=new_dataTrans)
```


```{r}
#Forward selection
intercept_mod <- lm(Life.expectancy~1, data=new_dataTrans)
add1(intercept_mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```

```{r}
#due to all value has same Pvalue. So Fvalues are compared and Schooling is chosen
mod <- lm(Life.expectancy~Schooling, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)

```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)

```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI+Polio, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI+Polio+Status, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)

```

```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI+Polio+Status+Measles, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```

```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI+Polio+Status+Measles+Hepatitis.B, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
```{r}
mod <- lm(Life.expectancy~Schooling+HIV.AIDS+Adult.Mortality+Diphtheria+thinness.5.9.years+
            Income.composition.of.resources+GDP+BMI+Polio+Status+Measles+Hepatitis.B+Population, data=new_dataTrans)
add1(mod, test="F",
     scope = ~Status+Adult.Mortality+HIV.AIDS+Diphtheria+
       BMI+Income.composition.of.resources+Polio+
       thinness.5.9.years+Population+Measles+Hepatitis.B+Alcohol+
       Total.expenditure+GDP+under.five.deaths+Schooling)
```
from the forward selection, it shows the result like above so,forward mod will be
```{r}
for_mod <- lm(Life.expectancy ~ Schooling + HIV.AIDS + Adult.Mortality + Diphtheria + 
                thinness.5.9.years + Income.composition.of.resources + GDP + 
                BMI + Polio + Status + Measles + Hepatitis.B + Population,data=new_dataTrans)
```


```{r}
#model with categorical data
category_mod <- lm(Life.expectancy ~ Schooling + HIV.AIDS + Adult.Mortality + Diphtheria + 
                thinness.5.9.years + Income.composition.of.resources + GDP + 
                BMI + Polio + Status + Measles + Hepatitis.B + Population,data=new_dataTrans)
```


```{r}
CP_PRESS <- function(model, sigma_full){
res <- resid(model)
hat_mod <- hatvalues(model)
CP <- sum(res^2)/sigma_full + 2*length(coef(model)) - length(res)
PRESS <- sum(res^2/(1-hat_mod)^2)
list(Cp=CP, PRESS=PRESS)
}
```

```{r}
sigma_q <- summary(int_mod)$sigma^2
size9_stat <- CP_PRESS(size_9, sigma_q)
size10_stat <- CP_PRESS(size_10, sigma_q)
size11_stat <- CP_PRESS(size_11, sigma_q)
sel_stat <- CP_PRESS(Step_mod, sigma_q)
```

```{r}
Life_frame <- data.frame(Model = c("Size_9","Size_10","Size_11","Stepwise"),p = c(8,9,10,13),
                          AdjR2 = c(summary(size_9)$adj.r.square,
                          summary(size_10)$adj.r.square,
                          summary(size_11)$adj.r.square,
                          summary(Step_mod)$adj.r.square),
                         CP = c(size9_stat$Cp,size10_stat$Cp,size11_stat$Cp,sel_stat$Cp),
                         PRESS = c(size9_stat$PRESS,size10_stat$PRESS,size11_stat$PRESS,sel_stat$PRESS))

Life_frame
```
Model with size8 shows the smallest Adjusted R-squared, then Stepwise model has the highest Adjusted R-squared value
From the list of model, It can be found that size_11 of model has Adjusted R-squared of 0.8136134, then second leaset CP and PRESS value,

```{r}
autoplot(size_9)
```

```{r}
autoplot(size_10, title = "Model of Size 10")
```
```{r}
autoplot(size_11)
```

```{r}
autoplot(Step_mod)
```
Tried the forward and backward


```{r}

bc2<- MASS::boxcox(size_10)
which.max(bc2$y)
lambda <- bc2$x[which.max(bc2$y)]
lambda

```



```{r}
cox_mod <- lm((((Life.expectancy)^1.959596)-1/1.959596) ~Adult.Mortality + BMI + Polio + Diphtheria+HIV.AIDS+
                GDP+thinness.5.9.years+Income.composition.of.resources+Schooling,data = new_dataTrans)
```


```{r}
autoplot(cox_mod)
```
```{r}
summary(cox_mod)
```
```{r}
summary(size_10)
```

```{r}
summary(size_10)
```

```{r}
category <- lm(Life.expectancy ~ Adult.Mortality + BMI + Polio + Diphtheria+HIV.AIDS+
                GDP+thinness.5.9.years+Income.composition.of.resources+Schooling+Diphtheria+
                 Adult.Mortality:Status+HIV.AIDS:Status+GDP:Status+Schooling:Status+BMI:Status,data=new_dataTrans)
```

```{r}
ggplot(new_dataTrans)+
  geom_point(aes(thinness.5.9.years,Life.expectancy, colour = Status))+
  theme_bw()
ggplot(new_dataTrans)+
  geom_point(aes(HIV.AIDS,Life.expectancy, colour = Status))+
  theme_bw()
ggplot(new_dataTrans)+
  geom_point(aes(Income.composition.of.resources,Life.expectancy, colour = Status))+
  theme_bw()
ggplot(new_dataTrans)+
  geom_point(aes(Schooling,Life.expectancy, colour = Status))+
  theme_bw()
```


```{r}
summary(category)
anova(category)
autoplot(category)
```




```{r}
CP_PRESS <- function(model, sigma_full){
res <- resid(model)
hat_mod <- hatvalues(model)
CP <- sum(res^2)/sigma_full + 2*length(coef(model)) - length(res)
PRESS <- sum(res^2/(1-hat_mod)^2)
list(Cp=CP, PRESS=PRESS)
}
```

```{r}
sigma_q <- summary(int_mod)$sigma^2
size9_stat <- CP_PRESS(size_9, sigma_q)
size10_stat <- CP_PRESS(size_10, sigma_q)
size11_stat <- CP_PRESS(size_11, sigma_q)
cox_stat <- CP_PRESS(cox_mod, sigma_q)
sel_stat <- CP_PRESS(Step_mod, sigma_q)
```


```{r}
sum(resid(size_9))
```

--Interpretation of the Model--
```{r}
summary(size_10)
```

```{r}
anova(size_10)

```

```{r}
confint(size_10, level=0.95)
```
`
```{r}
predict(size_10,new_dataTrans,interval = "confidence", level=0.95)
```




